---
title: "COVID-19 Infected Cases in Toronto"
author: "Jinwen Tao"
date: "2021-2-10"
output:
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE}
#install.packages("opendatatoronto")
#install.packages("dplyr")  
library(dplyr) 
library(ggplot2)
library(scales)
```


## Data

```{r, echo=FALSE, message=FALSE}
library(opendatatoronto)
COVID_19_Cases_in_Toronto_packages <- search_packages("COVID-19 Cases in Toronto")
COVID_19_Cases_in_Toronto_resources <- COVID_19_Cases_in_Toronto_packages %>%
  list_package_resources()

COVID_19_Cases_in_Toronto_statistics <-COVID_19_Cases_in_Toronto_resources %>%
  get_resource()
COVID_19_Cases_in_Toronto_data <- data.frame(COVID_19_Cases_in_Toronto_statistics)

COVID_19_GROUP <- COVID_19_Cases_in_Toronto_data %>%
  filter(Age.Group != "NaN") %>%
  select( Age_group = Age.Group,
          Outcome = Outcome
          )
head(COVID_19_GROUP, n=10)
```


## Graph 

```{r, echo=FALSE, results='hide',message=FALSE}
Agetable <- table(COVID_19_GROUP$Age_group)
Agetable
```

This shows an overview of the infected population.The group that aged between 20 to 29 has the largest infected number which reported 3049 infected cases. Moreover, approximately half of the infected people are older than 50 years old and people who are 19 years old or younger only have 1311 infected cases.

```{r, echo=FALSE, results="hide",warning=FALSE,message=FALSE}
COVID_19_DATA <- COVID_19_GROUP %>%
  group_by(Age_group, Outcome) %>%
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct))
COVID_19_DATA

ggplot(COVID_19_DATA, 
       aes(x = factor(Age_group,
                      levels = c("19 and younger",
                                 "20 to 29 Years","30 to 39 Years",
                                 "40 to 49 Years","50 to 59 Years", 
                                 "60 to 69 Years","70 to 79 Years",
                                 "80 to 89 Years","90 and older"),
                      labels = c("≤19", "20-29",
                                 "30-39", "40-49", "50-59", 
                                 "60-69", "70-79", "80-89", "90+")),
           y = pct,
           fill = factor(Outcome, 
                         levels = c("ACTIVE", "FATAL", "RESOLVED")
                         ))) + 
  
  
  geom_bar(stat = "identity", position = "fill") +
  scale_y_continuous(breaks = seq(0, 1, .2), 
                     label = percent) +
  geom_text(aes(label = lbl), 
            size = 3, 
            position = position_stack(vjust = 0.5))+
  scale_fill_brewer(palette = "Set2") +
  labs(y = "PERCENT", 
       fill = "CURRENT OUTCOME",
       x = "AGE",
       title = " COVID 19 INFECTION OUTCOME BY AGES",
       caption = "Source: Toronto Public Health, 2020.") +
  theme_minimal()


```

By plotting a bar chart, we divide the infected population in Toronto into three main categories based on their age groups. This chart clearly demonstrate the percentage of the current state in each age group that whether the cases are reported as fatal outcome, a resolved outcome or still an active outcome. The fatal outcomes are the death cases and it could be viewed in the graph that the old-aged groups who are older than 50 years old have a higher mortality rate than other groups. Specifically, the death cases in 80+ groups exceed 25% which indicate the COVID-19 has a more serious effect on the elderly. Comparing with the infected people who younger than 39 years old, the death rate almost reaches zero. Nevertheless, there are relatively higher rate of active outcome in the young groups which means around 10% of them are still infected with COVID-19 and not recover yet. For people who aged under 70 years old, the resolved cases have percentage over 85%. This shows that a large proportion of the age groups under 70 years old are reported as recovered or had no symptoms over 14 days and not currently hospitalized.


## Graph Discussion
There are some obvious relationship between the age and the outcomes of the infected people. As the ages increase, the death rate of people who infected with COVID-19 demonstrates an upward trend. The old-aged groups (over 50 years old) are the most dangerous population since half of the COVID-19 infected cases are old groups and their fatal rate are extremely higher than other groups. This means the old-aged people have greatest risk of infection and cause severe illness or even lead to death. On the contrary, COVID-19 has less effect on young-aged and mid-aged groups and they are more likely to recover after infection. More specially, the population that has the age under 70 years old has a greater chance to get recovered after being infected. 
In general, the risk of getting infected by COVID-19 and causing death is higher in elderly since they  are more vulnerable and have weaken immune systems. Although the recovery rates are higher in young and mid- aged groups, they have the largest infected population and they should also take protective measures to prevent themselves from the COVID-19 infection. 






